\include{template}

\usepackage{color}
\lstset{numbers=left}
\begin{document}

\printanswers

\title{575: Number Theory}
\maketitle



\begin{center}
    This exam has \numquestions\ questions for a total of \numpoints\
    points.You have {\bf 12 Hours} to complete this exam.
\end{center}

% 35 lectures, 12 hours, so 20.5ish points per lecture.

\section{Statements}

\begin{questions}

    \question[1] State the division algorithm
    
    \begin{solution}
        Given any integers $a$ and $b$, with $a > 0$, there exist
        unique integers $q$ and $r$ such that $b = qa + r$, $0 \le
        r < a.$
    \end{solution}

    \question[1] Define when two numbers $a$ and $b$ are relatively
    prime.

    \begin{solution}

        $a$ and $b$ are relatively prime if $(a, b) = 1$.

    \end{solution}

\end{questions}

\section{Short Answer}

\begin{questions}

    0903

    % Niven 1.2 Example 2
    \question Find integers $x$ and $y$ to satisfy
                  $$42823x + 6409y = 17$$

    \begin{solution}

        $$ \begin{array}{cc} 

        
        $$42823 \cdot 1 + 6409 \cdot 0 = 42823$$
        $$

    \end{solution}

\question

\end{questions}

\section{Free Response}

\begin{questions}

    % Lecture 1 part I
    \question[5] Prove that, for all positive integers $n$,
        $$1^3 + 2^3 + \cdots + n^3 = \frac{n^2(n+1)^2}{4}.$$
    \begin{solution}
        Base case: $1^3 = \frac{1^2(2)^2}{4} = 1$.

        Induction Step: 
        $$\sum_{k=1}^{n+1} k^3 = (n+1)^3 + \sum_{k=1}^n k^3$$
        $$  (n+1)^3 + \frac{n^2(n+1)^2}{4}
          = \frac{ 4(n+1)^3 + n^2(n+1)^2 }{4}$$
        $$ = \frac{ (n+1)^2 ( 4(n+1) + n^2 ) }{4}
           = \frac{ (n+1)^2 (n+2)^2 }{4}.$$        
    \end{solution}


    % Lecture 1 part II
    \question
    \begin{parts}
        % Niven/et. al. 1.2.9
        \part[3] Show that if $ac | bc$, then $a | b$.
        \begin{solution}
            The hypothesis implies that there is $m$ such that,
            $$bc = acm$$
            $$b = am$$
            i.e., $a | b$.
        \end{solution}
        % Niven/et. al. 1.2.10
        \part[3] Show that if $a | b$ and $c | d$ then $ac | bd$.
        \begin{solution}
            By hypothesis there is $m$ and $n$ such that
            $$b = am,\ d = cn$$
            $$bd = ac(mn)$$
            So $ac | bd.$
        \end{solution}
    \end{parts}

    % Lecture 1 part III
    \question (\totalpoints\ points)
    \begin{parts}
        % Niven/et. al. 1.2.11
        \part[8] Prove that $4 \nmid (n^2 + 2)$ for any integer $n$.
        \begin{solution}
            Using the division algorithm write $n = 4q + r$ where $0 \le r
            < 4.$ Then 
            $$n^2 + 2 = 16q^2 + 8qr + r^2 + 2$$
            $$ = 4(4q^2 + 2qr) + r^2 + 2.$$
            Examining possibilities for $r$ we see that the possibilities
            for $r^2 + 2$ are: 2, 3, 6 and 11. If $n^2 + 2$ were divisible
            by 4, then by the linear combination rule of divisbility, we'd
            have that $n^2 +2 - 4(4q^2 + 2qr) = r^2 + 2$ were divisible by
            4. However, by exhaustion 2, 3, 6, and 11 are not. 
        \end{solution}       
    \end{parts}      
\end{questions}

\section{Proofs of Theorems}

\begin{questions}

    \question[2] Prove that the greatest common divisor of two non-zero
    numbers exist.

    \begin{solution}
    
        Recall there is only a finite number of divisors of any
        nonzero integer. Thus there is a finite number of common
        divisors of two non-zero numbers. By the well-ordering
        principle, there is a greatest common divisor.

    \end{solution}

\end{questions}


\end{document}
