\include{template}

\usepackage{color}
\lstset{numbers=left}
\begin{document}

\printanswers

\title{Econ 105}
\maketitle



\begin{center}
    This exam has \numquestions\ questions for a total of \numpoints\
    points. You have {\bf 2 Hours} to complete this exam.
\end{center}

% Syllabus
%
% Microeconomics:
% -------------
%
% 1. scarcity, effiency, specialization
% 2. Supply and demand: 1 MC, 1 SA, 
% 3. Consumer demand: 1 MC,  1 SA, 
% 4. Supply: production, costs and profits: 3 MC, 
% 5. Perfect competition 1 MC, 1 SA
% 6. Imperfect competition
% 7. Factor prices, incomes, and income distribution
% 8. market efficiency and market failures
%
% Macroeconomics
% -------------
%
% 9. Macro overview
% 10. Measuring total output and inflation
% 11. Short-run AE model of output
% 12. Fiscal policy and government budgets
% 13. Money, interest rate, and monetary policy
% 14. AD and AS
% 15. Unemployment and inflation
% 16. Economic growth, producivity and supply-side economics

\begin{questions}

  \section{Multiple Choice}

  % 2
  \question[2] At the equilibrium price for a good:
  \begin{choices}
    \choice prices will remain unchanged, even if there is excess
    demand.
    \choice there may be excess demand for the good but not excess
    supply.
    \choice shifts in the supply or demand curves will not cause price
    changes.
    \CorrectChoice the quantity supplied of the good must equal the
    quantity demanded.
  \end{choices}

  % 3
  \question[2] A consumer will be maximizing utility at
  \begin{choices}
    \choice any intersection of the budget line and an indifference
    curve
    \choice any point on the highest indifference curve shown on the
    indifference diagram.
    \CorrectChoice the point where the indifference curve is tangent
    to the budget line
    \choice any point on the budget line
    \choice any point inside the budget line.
  \end{choices}

  % 4
  \question[2] The price of coal rose and the quantity sold
  fell. Which of the following is consistent with this observation
  (everything else being equal)?
  \begin{choices}
    \choice the price of oil rose
    \choice the productivity of coal miners rose
    \CorrectChoice more costly coal mining equipment was installed
    \choice the population increased
  \end{choices}

  % 4
  \question[2] The law of increasing (opportunity) costs explains why
  the production possibilities frontier  (PPF) is
  \begin{choices}
    \choice downward sloping
    \choice upward sloping
    \choice bowed inward
    \CorrectChoice bowed outward
    \choice undefined because no market will exist in this case.
  \end{choices}

  % 4
  \question[2] Marginal cost is the
  \begin{choices}
    \choice total cost divided by the number of units produced.
    \choice variable cost divided by the number of units produced
    \choice extra average cost of producing an additional unit of
    output
    \CorrectChoice extra cost of producing an additional unit of
    output
  \end{choices}

  % 5
  \question[2] Which of the following is not characteristic of a
  perfectly competitive industry?
  \begin{choices}
    \CorrectChoice Price is the primary decision variable of
    individual firms in the industry
    \choice There are many buyers and sellers of the industry product
    \choice There are no barriers to exit from or entry to the
    industry
    \choice The products of all firms in the industry are identical
    \choice None of the above.
  \end{choices}

  \section{Short Answer}

  % 3 
  \question[5] Economists give two reasons why most demand curves have
  a negative slope. Name and explain one of them.
  \begin{solution}
    There are the income and substition effects:

    As the price of a commodity falls the good becomes relatively
    cheaper compared to substitute commodities and thus become
    prefered so demand rises. This is called the substition effect.

    Also as the price falls the real buying power of a consumer rises,
    thus enabling the consumer to buy more of that good so demand
    rises. This is called the income effect
  \end{solution}

  % 2
  \question[5] How would you expect the price elasticity of Coca-Cola
  to compare to the price elasticity of soft drinks as a whole?

  \begin{solution}
    The price elasticity of a good is the percent change in quanity
    consumed over the percent change in price. Since Coke is one of
    many soft drink options, the price elasticity of demand for
    Coca-Cola is more negative than for soft drinks as a whole. As the
    price increases, the quantity demand drops off more because there
    are many substititues. If the price increased for soft drinks as a
    whole, there are fewer subsititues (hard drinks, coffee, etc.)
    which are not as complete substititues.
  \end{solution}

  % 5 ?
  \question (\totalpoints\ points) A few years ago a report aired on
  60 minutes which showed the unhealthy conditions in the American
  chicken industry, and suggested that consumers were risking
  salmonella infection each time they ate chicken.

  Using graphs and words answer the following questions about the
  impact of this report on the chicken industry and on the individual
  chicken grower. You should assume that the US industry producing
  chickens is a perfectly competitive industry, and had been in a
  state of long run equilibrium prior to the report.

  \begin{parts}
    
    \part[5] Explain what will happen to the industry demand curve
    and the equilibrium levels of price and output in the industry in
    the short run. What will happen to the profits earned by chicken
    growers in the short run?

    \begin{solution}
      In the short run the demand curve shifts left, meaning that at
      the current price point more chickens will be produced than
      bought. There will be a surplus of goods and profits will be
      negative. 
    \end{solution}

    \part[5] Explain what will happen to the number of chicken
    growers in the industry, the equilibrium levels of price and
    output in the industry, and the profits earned by chicken growers
    in the long run.

    \begin{solution}
      As Growers lose profits they will leave and the supply curve
      also shifts left. This causes prices to rise in the
      industry. Thus individual firms see profits rise from less than
      zero to 0 in the long-run. Industry price remains to close to
      the original price in the long run, but industry output falls
      even further than in short run due to exit from industry.
    \end{solution}

  \end{parts}

\end{questions}

\end{document}
